### 2020-02-18 ACM/abstract.tex
refactored the structure of the repo; added abstract for ACM

\documentclass[format=sigconf]{acmart}

\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
\usepackage{lipsum}

\title{Usage of N-gram models to improve autocompletion in Pharo}
\author{Myroslava Romaniuk}
\affiliation{\institution{Ukrainian Catholic University}}
\email{romaniuk@ucu.edu.ua}

\begin{document}

\begin{abstract}
  \par
  Code completion is one of the most essential features of any IDE.
  A good code completion tool can greatly improve the developer experience,
  whereas poor autocompletion can hinder the progress and discourage
  developers. Thus, it is important to find a way to make it as effective
  as possible. Specifically, in the case of Pharo, same name for a Smalltalk-like
  language and an open-source IDE, in the last year and a half a lot of work
  has been carried out to improve code completion. However, there are more
  potential improvements from which developers can only benefit: saving time
  and effort spent while coding remains an open question, and improving the
  relevance of suggestions and hence reducing time the developer takes to go
  through the list seem promising.

  In Pharo, the existing autocompletion implementation so far only allows for
  alphabetical sorting of results. While this approach is mostly usable, it
  completely ignores the potential of suggesting completion results based
  on code history and analysis, which is something that has been successfully
  adopted by other IDEs. Thus, in this work I intend to study how code
  completion in Pharo can be improved by using machine learning and statistical
  language models. In particular, the idea is to use the N-gram model, as it
  has been documented to be used for such a task, and, if successfully adapted
  to Pharo, might help enhance the code completion quality in general and the
  relevance of proposed completions in particular.

  There are many approaches to handling the process of completing code. The classic
  strategies have been to use lexical models (completing based on the prefix the user
  is typing) or to use semantic models (use code structure analysis information to
  tailor completions). In the latest years statistical, or machine learning, models
  have also gained popularity. For the most part when using ML, source code analysis
  is also often disregarded in lieu of simply training the model on the code base and
  inferring all the possible code dependencies from that.

  The famous paper "On The Naturalness of Software" (Hindle et al.) served as a sort
  of catalyst for the following research, applying statistical models to Pharo.
  There, as well as in “On the localness of software” (Tu et al.), “Code completion
  with statistical language models” (Raychev et al.), “Are deep neural networks the
  best choice for modeling source code?” (Hellendoorn et al.) and “Code completion
  with neural attention and pointer networks” (Li et al.), the structural information
  is not taken into consideration and the problem of improving code completion is
  reduced to a natural language processing problem, where predicting the next token
  is treated as predicting the next word in a sentence.

  The n-gram language model approach proposed by Hindle et al. is based on capturing
  high-level statistical regularity at the n-gram level by taking n-1 previous tokens
  that are already entered into the text buffer, and attempting to guess the next
  token. Using this model, it is possible to estimate the most probable sequences of
  tokens and suggest the most relevant code completions to developers.

  Therefore, in Pharo our idea is to have a sorting strategy based on an n-gram model,
  on top of the actual completion engine that works by analysing AST information of the
  source code, having actually utilised structural information. In this case, we are
  able to get semantically-appropriate results based on type whenever possible, and
  then sorting the results by relevance will help suggest much more accurate tokens for
  when the type is not known (as Pharo is a dynamically typed language).

  The goal of this will be to see:
  \begin{itemize}
    \item whether we can improve code completion in Pharo by sorting candidate
    completions with an n-gram language model
    \item whether we can build a tool based on a trained n-gram model that would
    propose completion fast enough to be used in an IDE
    \item how can we numerically evaluate the results of code completion produced
    by different completion strategies
  \end{itemize}
  \par
\end{abstract}

\maketitle

\end{document}