### 2020-02-19 ACM/abstract.tex
rewrote the abstract

\documentclass[format=sigconf]{acmart}

\usepackage{xcolor} %%% for comments
\usepackage{ifthen} %%% for comments
\usepackage{ulem} %%% for striked out text

% comments \nb{label}{color}{text}
\newboolean{showcomments}
\setboolean{showcomments}{true}
\ifthenelse{\boolean{showcomments}}
	{\newcommand{\nb}[3]{
		{\colorbox{#2}{\bfseries\sffamily\scriptsize\textcolor{white}{#1}}}
		{\textcolor{#2}{\sf\small$\blacktriangleright$\textit{#3}$\blacktriangleleft$}}}
	 \newcommand{\version}{\emph{\scriptsize$-$Id$-$}}}
	{\newcommand{\nb}[3]{}
	 \newcommand{\version}{}}
\newcommand{\md}[1]{\nb{Marcus}{red}{#1}}
\newcommand{\oz}[1]{\nb{Oleks}{olive}{#1}}
\newcommand{\mr}[1]{\nb{Myroslava}{teal}{#1}}

\title{Improving Code Completion in Pharo\\Using N-gram Language Models}
%\title{Usage of N-gram models to improve autocompletion in Pharo}
\author{Myroslava Romaniuk}
\affiliation{\institution{Ukrainian Catholic University}}
\email{romaniuk@ucu.edu.ua}

\begin{document}

\begin{abstract}
In this paper we present applying statistical language models to improve code completion
in Pharo. In particular, the goal is to use n-gram models for sorting the completion
candidates and in such a way increase the relevancy of untyped suggestions.

\section{Introduction}
\oz{Ask Stephane if you should add an affiliation with Inria and RMoD}
Code completion is one of the essential features of any IDE that greatly improves the
developer experience. Thus, it is important to find a way to make it as effective as
possible. Specifically, in the case of Pharo, a dynamically-typed programming language
and IDE, in the last year and a half a lot of work has been carried out to improve code
completion. However, there are more potential improvements from which developers can only
benefit, such as improving the relevance of suggestions and hence reducing time and
effort the developer spends to go through the list of completions.

\section{Context}
The current implementation of code completion in Pharo is based on abstract syntax tree
(AST) information of source code, where we are able to determine the structure of the code
and infer the type information where possible. Once we have a list of semantically suitable
completion candidates, they are sorted alphabetically. However, this approach ignores the
potential of suggesting completion results based on code history, which is something that
has been successfully adopted by other IDEs. In this work I intend to study how code
completion in Pharo can be improved using machine learning (ML) and statistical language
models. In particular, the idea is to use the N-gram model, as it has been documented to be
used for such a task \cite{Hind12a}, and, if successfully adapted to Pharo, might help
enhance the quality of code completion.

There are many approaches to handling the process of completing code. The classic strategies
use lexical \footnote{https://code.visualstudio.com/docs/editor/intellisense} (completing
based on the prefix the user is typing) or semantic (use code structure analysis information
to tailor completions) models. In the latest years statistical and ML models have also gained
popularity. For the most part when using ML, source code analysis is also often disregarded,
and the results of simply training the model on the code base and inferring all the possible
code dependencies are used instead.

\section{Related Works}
My work was inspired by the paper "On The Naturalness of Software" \cite{Hind12a}. There, as well
as in \cite{Tu14a,Rayc14a,Hell17a,Li17a}, the structural information is not taken into consideration
and the problem of improving code completion is reduced to a natural language processing problem,
where predicting the next token is treated as predicting the next word in a sentence.

The n-gram language model approach proposed by Hindle et al. \cite{Hind12a} is based on capturing statistical
regularity at the n-gram level by taking n-1 previous tokens that are already entered into the text
buffer, and attempting to guess the next token. Using this model, it is possible to estimate the
most probable sequences of tokens and suggest the most relevant code completions to developers.

\section{Our Approach}
What we want to do in Pharo is slightly different. Unlike the implementations of \cite{Tu14a,Hind12a,Rayc14a,Hell17a,Li17a},
where code completion is based solely on the results of statistical or machine learning models, the code
completion in Pharo works by making use of AST information and giving contextually relevant results.
However, as Pharo is a dynamically-typed language, type information is not always available, and the
suggested results, after being sorted alphabetically, require more effort to go through and manually
narrow down. This is where we think a sorting strategy based on an n-gram model will work better. In this
case tokens will be suggested according to their probability of appearing in the already typed sequence,
which will allow us to propose more accurate suggestions even without knowing the exact type of the token.

The estimated contributions of this work are the following:
\begin{itemize}
  \item improve code completion in Pharo by sorting candidate completions with an n-gram language model
  \item build a tool based on a trained n-gram model that would propose completion fast enough
  to be used in an IDE
\end{itemize}

\section{Future Work}
We would like to have a methodology to numerically evaluate the results of code completion
produced by different completion strategies. In particular, to have a robust estimation of
the improvement of various n-gram models compared to the alphabetical completion sorting
and among each other.

\section{Conclusion}
In this paper we proposed applying n-gram language models when sorting completion candidates
in Pharo. Having an AST-based completion engine, we are not always able to infer type information
from the nodes, and hope to compensate for it and improve the relevance of suggestions by
implementing a fast n-gram based sorting to be used in the Pharo IDE.

\end{abstract}

\maketitle

\bibliographystyle{ACM-Reference-Format}
\bibliography{rmod,others}

\end{document}