### 2020-05-26 IWST/paper.tex
adding the template for the paper

\documentclass[sigplan,screen]{acmart}
\usepackage{balance}

\begin{document}
\title{Title}
\author{Myroslava Romaniuk}
\affiliation{\institution{Ukrainian Catholic University} \city{Lviv} \country{Ukraine}}
\email{romaniuk@ucu.edu.ua}
\begin{abstract}
Abstract
\end{abstract}

\maketitle
\bibliographystyle{ACM-Reference-Format}
\bibliography{rmod,others}

\end{document}